### 2018-04-18 paper.Rmd
edit pass, tweaks throughout

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development, particularly in characterizing the early language learning environment.
  Access to these data can be complex for novices and difficult to automate for advanced users, however.
  To address these issues, we introduce *childes-db*, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API).
  Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

One of the key methods that has emerged to address this gap is the collection and annotation of speech to and by children, often in the context of the home. 
Starting with Roger Brown's [-@brown1973] work on Adam, Eve, and Sarah, audio recordings -- and more recently video recordings -- have been augmented with rich, searchable annotations that have allowed researchers to address a number of questions regarding the language learning environment.
Focusing on language learning in naturalistic contexts also reveals that children have in many cases productive and receptive abilities exceeding those revealed in experimental contexts; owing to challenges in lab-based experiments with young children, often their most revealing and sophisticated uses of language emerge in naturalistic play.

While corpora of early language acquisition are extremely useful, creating such corpora requires significant resources. 
Collecting and transcribing audio and video is costly and extremely time consuming -- even orthographic transcription (i.e., transcriptions with minimal phonetic detail) can take ten times the length of the original recording to transcribe [@macwhinney2000].
Automated, machine learning -based methods like automatic speech recognition (ASR) have provided only modest gains in efficiency, in that they are limited by the less-than-ideal acoustic properties of home recordings and the poor fit of language models built on adult-directed, adult-produced language samples for characterizing child-directed or child-produced speech.
Thus, researchers' desire for sufficient sample sizes and coverage in analyses of child language corpora can very quickly outstrip their resources. 
<!-- data pooling is necessary) -->

@@

Since its inception, a tremendous body of research has made use of CHILDES data.
Individual studies are too numerous to list, but classics include studies of morphological over-regularization [@marcus1992], distributional learning [@redington1998], word segmentation [@goldwater2009], the role of frequency in word learning [@goodman2008], and many others.
Some studies analyze individual examples in depth [e.g., @snyder2007], other track multiple child-caregiver dyads [e.g., @meylan2017], and still others use the aggregate properties of child or caregiver speech [e.g., @redingtonEtAl1998; @montagEtAl2015].
These are just a few of the thousands of studies that make use of CHILDES corpora.

Nonetheless, there are some outstanding challenges working with CHILDES, both for students and for advanced users.
The CHILDES ecosystem typically uses a specialized file format (CHAT), which is stored as plain text but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological, morphological, or syntactic structure of the utterance.
These files are usually analyzed using a command-line tool (CLAN), which has a number of specific tools that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU), and execute complex searches against the data.
While this system is flexible and powerful, mastering the CHAT codes and especially the CLAN tool with its many tools and flags can be daunting for a beginner or student.
These technical barriers decrease the ease of exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal.
CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- it can be tricky to build into a processing pipeline written in Python or R.
Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analyses typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

The practice of writing custom parsers is problematic for a number of reasons.
First, effort is wasted in implementing the same features again and again.
Second, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard.
Third, these parsing scripts are rarely shared or break with subsequent revisions to the dataset, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017 for an example].
Fourth, the CHILDES corpus itself is a moving target; computational work using the entire corpus at one time point may include a different set of data than subsequent work due to newly-contributed corpora or revised annotations of existing ones; currently, there is no simple way for researchers to document exactly what version of the corpus has been used short of creating a full mirror of the data.
These factors together lead to a lack of computational reproducibility, a major problem that keeps researchers from verifying or building on published research [@donoho2010;@stodden2016].

In the current manuscript, we describe a system for extending the functionality of CHILDES to address these issues.
Our system, childes-db, is a database-formatted mirror of CHILDES that allows programmatic access through an application programming interface (API). 
This infrastructure allows the creation of web applications for browsing and easily visualizing the data, facilitating classroom use of the dataset. 
Further, it can be accessed programmatically by advanced researchers, obviating the need to write one-off parsers of the CHAT format.
The database is versioned for access to previous releases, allowing computational reproducibility of particular analyses.

We begin by describing the architecture of childes-db and the web applications that we provide.
Next, we describe the `childesr` API, which provides a set of R functions for programmatic access to the data.
We conclude by presenting several worked examples of specific uses of the system--both web apps and the R API--for both research and teaching.

@@

```{r database-schema, fig.pos="h", out.width = "95%", fig.cap="Database schema for childes-db. Tokens are linked to superordinate groupings of utterances, transcripts, corpora, and collections (red arrows). All tokens and utterances are additionally associated with a participant (blue arrows)."}

@@

Many of the other tables in the database are hierarchical collections built out of tokens -- *utterance*, *transcript*, *corpus*, and *collection* -- which store attributes appropriate for each level of description.
Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.
An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number of morphemes (if the morphological structure is available in the original CHAT).
A *transcript* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES.
A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus.
Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES.
Because every record can be linked to a top-level collection (generally corresponding to a language), each table includes data from all languages represented in CHILDES.

Participants (e.g., children and caregivers) are represented separately from the token hierarchy because it is common for the same children to appear in multiple transcripts.
A participant identifier is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI = child, MOT = mother, FAT = father, etc.), and the range of ages (or age of corresponding child) for which they are observed.
For non-child participants (caregivers and others), the record additionally contains an identifier for the corresponding target child, such that data correpsonding to children and their caregivers can be easily associated.

@@

In order to import the CHILDES corpora into the MySQL schema described above, it must first be accurately parsed and subsequently vetted to ensure its integrity.
We parse the XML (eXtensible Markup Language) release of CHILDES hosted by \url{childes.talkbank.org} using the NLTK library in Python [@bird2004].
<!-- details of our contribution to the CHILDES corpus module in NLTK---didn't we change it? AS: I don't know if I would call it a "contribution" -- we just refactored it to make it more convenient for our purposes -->
Logic implemented in Python converts the linear, multi-tier parse into a tabular format appropriate for childes-db.
This logic includes decisions regarding what information sources are captured in the current release of the database, and which are left for future development (detailed below).

The data imported into childes-db is subject to data integrity checks to ensure that our import of the corpora is accurate and preferable over ad-hoc parsers developed by many individual researchers.
<!-- MCF: more testing details would be useful -->
In order to evaluate our success in replicating CLAN parses, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT.
In particular, the CLAN commands FREQ and MLU were used to compare total token counts and mean lengths of utterance for every speaker in every transcript; these numbers were compared to our own using the Pearson correlation coefficient. 
The results of the comparison were .99 and .98 for the unigram count and mlu data, respectively, indicating a reliable parse on our end.
<!-- SM: what were the results of these comparisons?  -->
<!-- AS: technically this is all we do right now SM: what more can we do for low technical investment-->

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of the time of writing, these changes are not systematically tracked.
We introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months or as warranted by changes in CHILDES.
By default, users interact with the most recent version of the database available. 
To support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific historical versions of the database.
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions AS: app functionality depends on db version in some cases, so nontrivial SM: let's punt-->

@@

Our web apps are built using Shiny, an R package that enables easy app construction using `R`.
Underneath the hood, each web app is making calls to our `childesr` API and subsequently plots the data using the popular R plotting package `ggplot2`.
A user's only task is to configure exactly what should be plotted through a series of buttons, sliders, and text boxes.
The user may specify what collection, corpus, child, age range, caregiver, etc., should be included in a given analysis.
The plot is displayed and updated in real-time, and the underlying data are also available for download alongside the plot to facilitate customization. All of these analyses may also be reproduced using the `childesr` package.

The web apps are intended for the casual user who seeks to easy extract developmental indices quickly and without any technical overhead.
We next walk through the three web apps provided on our website and describe their functionality in more detail.

@@

Many times a researcher will want to investigate the statistics of a given corpus (i.e the size, number of utterances, number of tokens, etc.) before choosing a target corpus for their analysis. This web app is intended to provide a basic overview of various corpora in the CHILDES ecosystem, facilitating familiarity with the different corpora and thus assisting with the overall analytic end-to-end process.
<!-- note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->

@@

# Using childes-db: Worked Examples
<!-- SM: we should note somewhere that examples all use Tidyverse, so that R users who don't know what %>% don't trip -->
In this section we give a number of examples of how childes-db can be used in both research and teaching, using both the web apps and the R API.

@@

From this figure, it seems likely that children hear "blue" more frequently early in development, but the trajectories of the "red" and "green" are less clear.
We also do not have a good sense of the errors of these measurements, are limited to just a few colors at a time before the plot becomes too busy, and cannot combine frequencies across speakers. <!-- SM:"busy" seems a bit informal -->
To perform this analysis in a more compelling and complete way, a user can use the `childesr` interface.

@@

For this analysis, we will define our words of interest as the basic color words of English (except for gray, which children hear very rarely).
We store these in the `colors` variable, and then use the `get_types` function from `childesr` to get the type frequency of each of these words in all of the corpora in CHILDES.
For demonstration, we look only at the types produced by the speakers in each corpus tagged as Mother and Father.
We also restrict ourselves to children from 1-5 years old (12 to 60 months), and look only at the American English corpora.

@@

Figure \ref{fig:color-plot} shows the results of these analyses: Input frequency varies significantly over the 1-5 year range for nearly every color word.

@@

Gender differences in language production have been found in multiple studies, notably in @eriksson2012differences, a large-scale study that spanned ten languages and found that females have a higher vocabulary than males in early childhood.
A researcher or student in child language acquisition may now investigate this difference with relative ease using `childesr`.
By pulling data from the `transcript_by_speaker` table, a user has access to the the multitude of derived linguistic measures that are often used to evaluate a child's grammatical development.
In this worked example we walk through a sample analysis that probes the gender differences in early grammatical development.

@@

This call pulls data from all collections and corpora, including those languages for which there are very sparse data.
In order to make any substantial inferences from our analysis, we begin by filtering languages for which there are a significant number of children. 
We also filter for children under the age of four years.

```{r echo=TRUE}
number_of_transcripts_threshold <- 575
max_age <- 4

included_languages <- stats %>%
  filter(target_child_age < max_age * 365) %>%
  group_by(language) %>%
  summarise(n = n()) %>%
  filter(n > number_of_transcripts_threshold) %>%
  .$language
```

Our `transcript_by_speaker` table contains multiple derived measures of language productivity and lexical diversity -- here we used MTLD [@mccarthy2005assessment], a measure that is more robust to transcript length, as a way of evaluating children's linguistic productivity.
We start by filtering those children for which a sex was defined, who speak a language in our subset, and are in the appropriate age range.
We then average their MTLD score across transcripts after having grouped the data by child, age, sex, and language.

@@

```{r echo=TRUE, warning=FALSE}
ggplot(data, aes(x = age_years, y = measure, 
                 color = target_child_sex)) +
  geom_point(alpha = .3) + 
  geom_line(data = mod_data, 
            aes(x = age_years, y = measure, color = target_child_sex)) +
  facet_wrap(~language) + 
  ggthemes::theme_few() + 
  ggthemes::scale_color_ptol(name = "Sex") + 
  xlim(0,max_age) + 
  ylim(0,15) + 
  xlab("Age (years)") + 
  ylab("MTLD") + 
  theme(legend.position = "bottom")
```

This plot reveals a slight gender difference in linguistic productivity in young children (displaying a moderate female advantage), replicating the results of @eriksson2012differences.
The goal of this analysis was to showcase an example of the the steps involved in using `childesr` for exploring the CHILDES dataset, as well as some of the potential hazards in working with a diverse set of corpora that were collectd through various experimental paradigms.

@@

One goal for courses on applied natural language processing (NLP) is for students to get hands-on experience using NLP tools to analyze real-world language data.
A primary challenge for the instructor is to decide how much time should be spent teaching the requisite programming skills for accessing and formatting language data, which are typically unstructured.
One pedagogical strategy is to abstract away these details and avoid having students deal with obtaining data and formatting text.
This approach shifts students' effort away from data cleaning and towards programming analyses that encourage the exploration and testing of interesting hypotheses.

The `childesr` API provides instructors with an easy-to-learn method for giving students programmatic access to child language data.
The aim is to reduce time spent getting data out of CHILDES and allow more course time for thinking about interesting questions that could be answered with NLP and child language data. <!-- SM: somewhat redundant with the above -->
Moreover, by writing our API in the R programming language, students will gain access to a general workflow can be integrated into an analysis pipeline that includes state of the art NLP tools. <!-- SM: sorry, what tools? there's very little NLP support in R b/c it's string handling is slow -->

>>>>

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development, particularly in characterizing the early language learning environment.
  Access to these data can be both complex for novices and also difficult to automate for advanced users, however.
  To address these issues, we introduce *childes-db*, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API).
  Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

One of the key methods that has emerged to address this gap is the collection and annotation of speech to and by children, often in the context of the home. 
Starting with Roger Brown's [-@brown1973] work on Adam, Eve, and Sarah, audio recordings -- and more recently video recordings -- have been augmented with rich, searchable annotations that have allowed researchers to address a number of questions regarding the language learning environment.
Focusing on language learning in naturalistic contexts also reveals that children have in many cases productive and receptive abilities exceeding those demonstrated in experimental contexts; owing to challenges in lab-based experiments with young children, often their most revealing and sophisticated uses of language emerge in naturalistic play.

While corpora of early language acquisition are extremely useful, creating such corpora requires significant resources. 
Collecting and transcribing audio and video is costly and extremely time consuming -- even orthographic transcription (i.e., transcriptions with minimal phonetic detail) can take ten times the length of the original recording to transcribe [@macwhinney2000].
Automated, machine learning-based methods like automatic speech recognition (ASR) have provided only modest gains in efficiency. 
Such systems are limited both by the less-than-ideal acoustic properties of home recordings and also by the poor fit of language models built on adult-directed, adult-produced language samples to child-directed and child-produced speech.
Thus, researchers' desire for data in analyses of child language corpora can very quickly outstrip their resources. 
<!-- data pooling is necessary) -->

@@

Since its inception, a tremendous body of research has made use of CHILDES data.
Individual studies are too numerous to list, but classics include studies of morphological over-regularization [@marcus1992], distributional learning [@redington1998], word segmentation [@goldwater2009], the role of frequency in word learning [@goodman2008], and many others.
Some studies analyze individual examples in depth [e.g., @snyder2007], other track multiple child-caregiver dyads [e.g., @meylan2017], and still others use the aggregate properties of child or caregiver speech [e.g., @redingtonEtAl1998; @montagEtAl2015].
<!-- These are just a few of the thousands of studies that make use of CHILDES corpora. -->

Nonetheless, there are some outstanding challenges working with CHILDES, both for students and for advanced users.
The CHILDES ecosystem uses a specialized file format (CHAT), which is stored as plain text but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological, morphological, or syntactic structure of the utterance.
These files are usually analyzed using a command-line program (CLAN) that allows users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU), and execute complex searches against the data.
While this system is flexible and powerful, mastering the CHAT codes and especially the CLAN tool with its many functions and flags can be daunting.
These technical barriers decrease the ease of exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal.
CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- it can be tricky to build into a processing pipeline written in Python or R.
Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analysis typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

The practice of writing custom parsers is problematic for a number of reasons.
First, effort is wasted in implementing the same features again and again.
Second, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard.
Third, these parsing scripts are rarely shared -- and when when they are, they typically break with subsequent revisions to the dataset -- leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017 for an example].
Fourth, the CHILDES corpus itself is a moving target; computational work using the entire corpus at one time point may include a different set of data than subsequent work due to newly-contributed corpora or revised annotations of existing ones. Currently, there is no simple way for researchers to document exactly what version of the corpus has been used short of creating a full mirror of the data.
These factors together lead to a lack of *computational reproducibility*, a major problem that keeps researchers from verifying or building on published research [@donoho2010;@stodden2016].

In the current manuscript, we describe a system for extending the functionality of CHILDES to address these issues.
Our system, childes-db, is a database-formatted mirror of CHILDES that allows access through an application programming interface (API). 
This infrastructure allows the creation of web applications for browsing and easily visualizing the data, facilitating classroom use of the dataset. 
Further, the database can be accessed programmatically by advanced researchers, obviating the need to write one-off parsers of the CHAT format.
The database is versioned for access to previous releases, allowing computational reproducibility of particular analyses.

We begin by describing the architecture of childes-db and the web applications that we provide.
Next, we describe the `childesr` API, which provides a set of R functions for programmatic access to the data.
We conclude by presenting several worked examples of specific uses of the system--both web apps and the R API--for research and teaching.

@@

```{r database-schema, fig.pos="t", out.width = "95%", fig.cap="Database schema for childes-db. Tokens are linked to superordinate groupings of utterances, transcripts, corpora, and collections (red arrows). All tokens and utterances are additionally associated with a participant (blue arrows)."}

@@

Many of the other tables in the database are hierarchical collections built out of tokens -- *utterance*, *transcript*, *corpus*, and *collection* -- which store attributes appropriate for each level of description.
Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.
An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number of morphemes (if the morphological structure is available in the original CHAT).
A *transcript* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES.
A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpora.
Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES.
Because every record can be linked to a top-level collection (generally corresponding to a language), each table includes data from all languages represented in CHILDES.

Participants (e.g., children and caregivers) are represented separately from the token hierarchy because it is common for the same children to appear in multiple transcripts.
A participant identifier is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI = child, MOT = mother, FAT = father, etc.), and the range of ages (or age of corresponding child) for which they are observed.
For non-child participants (caregivers and others), the record additionally contains an identifier for the corresponding target child, such that data corresponding to children and their caregivers can be easily associated.

@@

In order to import the CHILDES corpora into the MySQL schema described above, it must first be accurately parsed and subsequently vetted to ensure its integrity.
We parse the XML (eXtensible Markup Language) release of CHILDES hosted by \url{childes.talkbank.org} using the NLTK library in Python [@bird2004].
<!-- details of our contribution to the CHILDES corpus module in NLTK---didn't we change it? AS: I don't know if I would call it a "contribution" -- we just refactored it to make it more convenient for our purposes -->
Logic implemented in Python converts the linear, multi-tier parse into a tabular format appropriate for childes-db.
This logic includes decisions that we review below regarding what information sources are captured in the current release of the database and which are left for future development.

The data imported into childes-db is subject to data integrity checks to ensure that our import of the corpora is accurate and preferable over ad-hoc parsers developed by many individual researchers.
<!-- MCF: more testing details would be useful -->
In order to evaluate our success in replicating CLAN parses, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT.
We used the CLAN commands FREQ and MLU to compare total token counts and mean lengths of utterance for every speaker in every transcript and compared these these values to our own using the Pearson correlation coefficient. 
The results of the comparison were .99 and .98 for the unigram count and MLU data, respectively, indicating reliable parsing.
<!-- SM: what were the results of these comparisons?  -->
<!-- AS: technically this is all we do right now SM: what more can we do for low technical investment-->

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of time of writing, these changes are not systematically tracked.
To facilitate reproducibility of past analyses, we introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months or as warranted by changes in CHILDES.
By default, users interact with the most recent version of the database available. 
To support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific historical versions of the database.
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions AS: app functionality depends on db version in some cases, so nontrivial SM: let's punt-->

@@

Our web apps are built using Shiny, an R package that enables easy app construction using `R`.
Underneath the hood, each web app is making calls to our `childesr` API and subsequently plots the data using the popular R plotting package `ggplot2`.
A user's only task is to configure exactly what should be plotted through a series of buttons, sliders, and text boxes.
The user may specify what collection, corpus, child, age range, caregiver, etc., should be included in a given analysis.
The plot is displayed and updated in real-time, and the underlying data are also available for download alongside the plot to facilitate customization. 
All of these analyses may also be reproduced using the `childesr` package, but the web apps are intended for the casual user who seeks to easy extract developmental indices quickly and without any technical overhead.
<!-- We next walk through the three web apps provided on our website and describe their functionality in more detail.  -->

@@

Many times a researcher will want to investigate the statistics of corpora (e.g., their size, number of utterances, number of tokens) before choosing a target corpus or set of corpora for a project.
This web app is intended to provide a basic overview of various corpora in the CHILDES ecosystem, facilitating familiarity with the different corpora and thus assisting with the overall analytic end-to-end process.
<!-- note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->

@@

# Using childes-db: Worked Examples

In this section we give a number of examples of how childes-db can be used in both research and teaching, using both the web apps and the R API. Note that all of these examples use `dplyr` syntax [@R-dplyr]; several accessible introductions to this framework are available online [e.g., @wickham2016].

@@

From this figure, it seems likely that children hear "blue" more frequently early in development, but the trajectories of the "red" and "green" are less clear.
We also do not have a good sense of the errors of these measurements, are limited to just a few colors at a time before the plot becomes too busy, and cannot combine frequencies across speakers. <!-- SM:"busy" seems a bit informal; MCF: I'm ok with it. :) -->
To perform this analysis in a more compelling and complete way, a user can use the `childesr` interface.

@@

For this analysis, we will define our words of interest as the basic color words of English (except for gray, which children hear very rarely).
We store these in the `colors` variable, and then use the `get_types` function from `childesr` to get the type frequency of each of these words in all of the corpora in CHILDES.
For demonstration, we look only at the types produced by the speakers in each corpus tagged as Mother and Father.
We also restrict ourselves to children from 1--5 years old (12--60 months), and look only at the American English corpora.

@@

Figure \ref{fig:color-plot} shows the results of these analyses: Input frequency varies substantially over the 1--5 year range for nearly every color word.

@@

Gender differences in language production are well-documented. For example, @eriksson2012differences used parent-report data from ten languages and found that female children had larger vocabularies on average than male children in nearly every language.
Comparable cross-linguistic analysis of naturalistic production data has not been conducted, however, and these differences are easy to explore using `childesr`.
By pulling data from the `transcript_by_speaker` table, users can access derived linguistic measures.
In this worked example we walk through a sample analysis that probes the gender differences in early lexical diversity.

@@

This command retrieves data from all collections and corpora, including those languages for which there are very sparse data.
In order to make any substantial inferences from our analysis, we begin by filtering languages for which there are a large number of transcripts ($>500$). 
We also restrict our analysis to children under the age of four years.

```{r echo=TRUE}
number_of_transcripts_threshold <- 500
max_age <- 4

included_languages <- stats %>%
  filter(target_child_age < max_age * 365) %>%
  group_by(language) %>%
  summarise(n = n()) %>%
  filter(n > number_of_transcripts_threshold) %>%
  pull(language)
```

Our `transcript_by_speaker` table contains multiple derived measures of lexical diversity -- here we used MTLD [@mccarthy2005assessment], a measure that is relatively robust to transcript length.
We start by filtering those children for which a sex was defined, who speak a language in our subset, and are in the appropriate age range.
We then average their MTLD score across transcripts after having grouped the data by child, age, sex, and language.

@@

```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x = age_years, y = measure, 
                 color = target_child_sex)) +
  geom_point(alpha = .3) + 
  geom_line(data = mod_data, 
            aes(x = age_years, y = measure, color = target_child_sex)) +
  facet_wrap(~language) + 
  ggthemes::theme_few() + 
  ggthemes::scale_color_ptol(name = "Sex") + 
  xlim(0,max_age) + 
  ylim(0,15) + 
  xlab("Age (years)") + 
  ylab("MTLD") + 
  theme(legend.position = "bottom")
```

This plot reveals a slight gender difference in linguistic productivity in young children (displaying a moderate female advantage), replicating the results of @eriksson2012differences.
The goal of this analysis was to showcase an example of the the steps involved in using `childesr` for exploring the CHILDES dataset, as well as some of the potential hazards in working with a diverse set of corpora that were collected through various experimental paradigms.

@@

One goal for courses on applied natural language processing (NLP) is for students to get hands-on experience using NLP tools to analyze real-world language data.
A primary challenge for the instructor is to decide how much time should be spent teaching the requisite programming skills for accessing and formatting language data, which are typically unstructured.
One pedagogical strategy is to abstract away these details and avoid having students deal with obtaining data and formatting text.
This approach shifts students' effort away from data cleaning and towards programming analyses that encourage the exploration and testing of interesting hypotheses. In particular, the `childesr` API provides instructors with an easy-to-learn method for giving students programmatic access to child language data.

<!-- The aim is to reduce time spent getting data out of CHILDES and allow more course time for thinking about interesting questions that could be answered with NLP and child language data. <!-- SM: somewhat redundant with the above --> -->
<!-- Moreover, by writing our API in the R programming language, students will gain access to a general workflow can be integrated into an analysis pipeline that includes state of the art NLP tools. <!-- SM: sorry, what tools? there's very little NLP support in R b/c it's string handling is slow.  --> -->
